% ===========================================================================
% Related Work content (section header defined in main.tex)
% ===========================================================================

\subsection{Rule-Based I/O Analysis}
HPC I/O diagnosis has traditionally relied on expert-crafted heuristics.
Drishti~\cite{bez2022drishti} applies 30 threshold rules to Darshan logs~\cite{snyder2016darshan},
flagging issues such as small I/O operations or access misalignment when
counters exceed fixed cutoffs (e.g., $>$10\% small operations).
WisIO~\cite{devarajan2025wisio} extends this approach with over 800 rules
organized across process, file, and timeline perspectives, achieving
340K detections per second.
While both tools are practical and well-engineered, rule-based systems
scale linearly with pattern complexity: each new bottleneck type requires
a human expert to encode a new rule.
They also evaluate rules independently, missing cross-feature interactions
(e.g., small I/O combined with random access patterns).
On our benchmark-derived ground-truth test set, Drishti achieves
Micro-F1\,=\,0.384 and WisIO achieves 0.318, compared to 0.923 for IOSage's
learned classifier, which captures nonlinear feature interactions
without manual rule authoring.

\subsection{ML for I/O Performance Analysis}
Several works apply machine learning to Darshan-derived features.
AIIO~\cite{dong2023aiio} trains an ensemble of five models (XGBoost,
LightGBM, CatBoost, MLP, TabNet) to predict aggregate I/O throughput from
POSIX and Lustre counters, using SHAP to identify influential features.
Bang et al.~\cite{bang2020clustering} cluster HPC jobs by I/O behavior using
unsupervised learning on Darshan counters, while
Paul et al.~\cite{paul2022characterizing} characterize ML workloads on Summit
through keyword-based job classification.
These studies address either throughput prediction or workload characterization,
but none formulates I/O diagnosis as multi-label classification
across distinct bottleneck dimensions.
AIIO's regression target (bytes/time) indicates \emph{how slow} I/O is
but not \emph{why}; its SHAP attributions show features correlated with
throughput rather than with specific bottleneck types.
IOSage reformulates the problem as 8-dimensional multi-label classification
with per-label SHAP attributions and pairs detection with
LLM-generated code-level fix recommendations.

\subsection{LLM-Based I/O Diagnosis}
Recent work applies large language models directly to I/O analysis.
ION~\cite{egersdoerfer2024ion} demonstrated a proof-of-concept using LLM
in-context learning for Darshan log interpretation, though it reported
no quantitative accuracy metrics.
Its successor IOAgent~\cite{egersdoerfer2025ioagent}, released as
IONavigator, adds RAG retrieval from 66 research papers and tree-based
merging of per-module diagnoses.
On 9 real-application traces from TraceBench, IONavigator produces 31 false
positive diagnoses against 15 true positives (Precision\,=\,0.326),
reflecting LLM overgeneration: when given I/O profiles, LLMs tend to discuss
all possible bottleneck categories rather than identifying only those present.
STELLAR~\cite{egersdoerfer2025stellar} applies multi-agent LLM reasoning
to Lustre file system parameter tuning, achieving up to 7.8$\times$ speedup;
its authors explicitly document LLM hallucination on storage parameters
(Figure~2 in their paper).
KORAL~\cite{akewar2026koral} combines knowledge graphs with LLMs for SSD
operational analysis, validating the trend toward structured knowledge
grounding.
IOSage addresses the overgeneration problem by using an ML classifier as a
precision gate: the LLM receives only dimensions where the classifier
exceeds a confidence threshold, reducing false positives from 31 to 4
(87\% reduction) on the same TraceBench traces.

\subsection{ML+LLM Hybrid Systems}
The pattern of combining ML detection with LLM reasoning has been validated
outside the I/O domain.
RCACopilot~\cite{chen2024rcacopilot} uses ML-based incident matching to
provide demonstrations for GPT-4 root cause analysis of cloud outages;
without its ML component, LLM accuracy drops from
Micro-F1\,=\,0.766 to 0.026.
The OpenRCA benchmark~\cite{openrca2025} reports that the best LLM-only agent
achieves 11.34\% accuracy on software root cause analysis, reinforcing
that LLMs alone are insufficient for precise diagnostic tasks.
IOSage applies this validated pattern to HPC I/O for the first time,
combining a 0.923 Micro-F1 multi-label classifier with a benchmark-grounded
knowledge base of 623 configurations that maps Darshan signatures to
validated code fixes with measured speedups.
